---
title: "Stock Market Prediction Project"
output:
  html_notebook: default
  pdf_document: default
---

Our goal is to create insights and draw predictions for the percentage returns from a stock market dataset which consists of percentage returns for the S&P 500 stock index over 1250 days in the past, from 2001 until the end of 2005. 

For each date, we have recorded 

1. Percentage Returns for each of the five previous trading days, Lag1 to lag5
2. Volumn: the number of shares traded on the previous day, in billions
3. Today: the percentage return on the date in question
4. Direction: whether the market was Up or Down on this date

```{r}
library(ISLR)
names(Smarket)
```

```{r}
dim(Smarket)
```


```{r}
summary(Smarket)
```
Notice the max and min for today, check for any outliers. 

Next, we would like to produce a matrix that contains all of the pairwise correlations among the predictors in a data set. (Notice here we drop the variable Direction since it is qualitative.)
```{r}
cor(Smarket[,-9]) #everthing but the 9th column
```
As we can see, the correlations between the lag variables and today's returns are close to zero. In other words, there apprears to be little correlations between today's returns and previous days' returns. The only substantial correlation is between Year and Volumn. 
```{r}
attach(Smarket) #access the variable without having the reference
plot(Volume)
```
By plotting the data, we see that Volumn is increasing over time. In other words, the average number of shares traded daily increased from 2001 to 2005. 

#### Logistic Regression
Next, we want to predit Direction using lag1 through lag 5 and Volume by fitting a logistic regression model. 
```{r}
glm.fits=glm(Direction~Lag1+Lag2+Lag3+Lag4+Lag5+Volume, family=binomial, data=Smarket) #famil=binomial tells R that we want a logistics regression model
summary(glm.fits)
```
The smallest p-value here is associated with Lag1. The negative coefficient fro this predictor suggests that if the market had a positive return yesterday, then it is more likely to go down today. However, at a value of 0.15, the p-value is still considered large, and so there is no clear evidence of a real relationship between Lag1 and Direction.

```{r}
coef(glm.fits)
```

```{r}
print(coef(glm.fits)[5]) #handy way to create report
```

```{r}
summary(glm.fits)$coef
```

```{r}
summary(glm.fits)$coef[,4]
```

The predict() function can be used to predict the probability that the market will go up, given values of the predictors. 
```{r}
glm.probs=predict(glm.fits,type='response')
#The type='response' tells R to ouput the probabilities of the form P(Y=1|X) instead of logit.
glm.probs[1:10] #The first 10 probability of the market going up. 
```

```{r}
contrasts(Direction) # To know 1 is for Up. How R interpret the direction up or down. 
```

```{r}
glm.pred=rep("Down", 1250)
glm.pred[glm.probs>.5]='up'
glm.pred[1:20]
```
```{r}
#confusion matrix
table(glm.pred, Direction)
# training data error 
```

```{r}
mean(glm.pred==Direction)
```
```{r}
train=(Year<2005)
Smarket.2005=Smarket[!train, ] #!train is to negate boolean 
dim(Smarket.2005)
```

```{r}
Direction.2005=Direction[!train]
glm.fits=glm(Direction~Lag1+Lag2+Lag3+Lag4+Lag5+Volume, data=Smarket, subset=train,family=binomial)
glm.probs=predict(glm.fits,Smarket.2005,type='response')
glm.probs[1:15]
```
```{r}
glm.probs[1:15]
glm.pred=rep("Down",252)
glm.pred[glm.probs>.5]="Up"
table(glm.pred,Direction.2005)
mean(glm.pred==Direction.2005) #not particilarly good
```
```{r}
train=(Year<2005)
Smarket.2005=Smarket[!train, ] #!train is to negate boolean 
dim(Smarket.2005)
Direction.2005=Direction[!train]
glm.fits=glm(Direction~Lag1+Lag2, data=Smarket, subset=train,family=binomial)
glm.probs=predict(glm.fits,Smarket.2005,type='response')
glm.probs[1:15]
glm.pred=rep("Down",252)
glm.pred[glm.probs>.5]="Up"
table(glm.pred,Direction.2005)
mean(glm.pred==Direction.2005) #improving 
```


